\thetitle

### 2010-03-22 ITP10/hybrid-12oclock.tex
Add ITP10 paper sources.
Ignore-this: 49de24983449dd9b9daef53ef1da6000

darcs-hash:20100322174102-bab43-dcc3cc0af50eb50c30ef52b815fff99223eb4669.gz

\documentclass[runningheads]{llncs}

\newcommand{\eelis}{Eelis van der Weegen}
\newcommand{\herman}{Herman Geuvers}
\newcommand{\dan}{Dan Synek}
\newcommand{\adam}{Adam Koprowski}

\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}

\usepackage{color}
\usepackage{amsmath}
\usepackage[colorlinks,pdftitle=\thetitle,pdfauthor=]{hyperref}

@@

%include polycode.fmt

%format pipe = "\mathopen{|}"

%format times = "\times"

%format . = "."
%format forall = "\forall"
%format Next = "\mathbf{Next}"
%format Obligation = "\mathbf{Obligation}"
%format Class = "\mathbf{Class}"
%format Fixpoint = "\mathbf{Fixpoint}"
%format Definition = "\mathbf{Definition}"
%format Eval = "\mathbf{Eval}"
%format Notation = "\mathbf{Notation}"
%format Lemma = "\mathbf{Lemma}"
%format Proof = "\mathbf{Proof}"
%format Qed = "\mathbf{Qed}"
%format Defined = "\mathbf{Defined}"
%format Function = "\mathbf{Function}"
%format Proposition = "\mathbf{Proposition}"
%format Record = "\mathbf{Record}"
%format Program = "\mathbf{Program}"
%format Inductive = "\mathbf{Inductive}"
%format Theorem = "\mathbf{Theorem}"
%format Let = "\mathbf{Let}"
%format Instance = "\mathbf{Instance}"
%format Variable = "\mathbf{Variable}"
%format Variables = "\mathbf{Variables}"
%format measure = "\mathbf{measure}"

%format prod a b = a "\times " b
%format sp = "\ "
%format frac a b = "\frac{" a "}{" b "}"

%format == = "= "
%format </ = "\nless "
%format /= = "\ne "
  % Without these, lhs2tex uses three-line equal signs.

%format fcmp = "\circ"
%format match = "\mathbf{match}"
%format with = "\mathbf{with}"
%format end = "\mathbf{end}"
%format fun = "\lambda\hspace{-1mm} "
%format nat = "\mathbb{N}"
%format CRplain = "\texttt{CR}"
%format CR = "\mathbb{R}"
%format CRasCSetoid = CR
%format exists = "\exists "
%format log2ceil = "\log_{2}\hspace{-0.5mm}"
%format over = "\mathbf{over}"
%format INR (a) = a
%format `union` = "\cup "
%format /\ = "\wedge "
%format ~ = "\neg "
%format \/ = "\vee"
%format <-> = "\leftrightarrow "
%format ==> = "\Longrightarrow "
%format `subsetof` = "\subseteq "
%format phi = "\phi "
%format False = "\bot "

%format sp1mm = "\hspace{1mm} "

%%format { = "\hspace{0.9mm}\{"
  % so that parameters declared with {...} are spaced like those declared with (...)
    % disabled because it makes record defs ugly

%format (pow (a) (b)) = a"^"b

%format In (a) (b) = a "\in " b
%format `NotIn` = "\notin"

@@

\arrayhs % i.e. no page breaks in code blocks

@@

\author{\eelis}
\institute{Institute for Computing and Information Sciences\\ Raboud University Nijmegen}

@@

\begin{abstract}
This technical report documents our development of a hybrid system safety prover, implemented in Coq using the abstraction method introduced by Alur in \cite{alur}. The development includes: a formalization of the structure of hybrid systems; a systematic approach and generic set of support utilities for the construction of an abstract system (consisting of decidable ``overestimators'' of abstract transitions and initiality) faithfully representing a (concrete) hybrid system; a translation of abstract systems to graphs enabling decision of abstract state reachability using a certified graph reachability algorithm; a proof of an example hybrid system (taken from \cite{alur}) generated using this tool stack. The development critically relies on the computable real number implementation part of the C-CoRN library of formalized constructive mathematics. \todo{Hm, anything else worth mentioning?}\footnote{This research was supported by the BRICKS/FOCUS project 642.000.501 ``Advancing the Real use of Proof Assistants''.}

@@

\setcounter{tocdepth}{2}

\tableofcontents
\todo{How do I get rid of that silly author index at the top of the TOC?!}

@@

In \cite{alur}, Alur presents a method for automated hybrid system safety verification based on the construction of an \emph{abstract} hybrid system (essentially a finite automaton) corresponding to the hybrid system of interest. The abstract system is constructed such that traces in the original system are represented in the abstract system. Consequently, one can draw conclusions about reachability of states in the concrete system from analysis of state reachability in the abstract system. Since the abstract system is an entirely finite discrete object (unlike the concrete system), reachability can simply be computed (using any graph reachability algorithm). Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it is amenable to what amounts to brute force.

The prototype implementation described in \cite{alur} was developed in a conventional programming language using libraries that use ordinary floating point arithmetic \todo{which language/libraries exactly?}. The potential for bugs and floating point artifacts inherent in this approach limits the confidence one can justifiably have in safety judgements made by such an implementation. In this report, we describe a reimplementation of the basic technique in Coq, a proof assistant based on a rich type theory that also functions as a programming language, letting one develop ``certified'' programs: programs whose correctness is proved formally within the system. In our case, the program will be one that produces machine-checked proofs of hybrid system safety, obtained using the abstraction method.

Our development relies crucially on computation using real numbers, for which we use the computable real number implementation developed by Russell O'Connor \cite{something}, and part of the C-CoRN library of formalized constructive mathematics \cite{something}. Indeed, this development showcases its use in a concrete and practical application, and subtleties regarding the nature of these computable reals are reflected in this development in several ways, which we will discuss in some detail.

@@

 % In our case, the specification of the program we write is this: given a description of a concrete system, the program  . The whole implementation, including the computational part, serves only to prove a single theorem. Hence, one essentially only needs to trust Coq to be convinced of the safety proofs created by our development.

@@

%Wherever we say ``|CR|'', we are really referring to the type |CRplain| of computable reals, part of C-CoRN.

%The computational steps become part of the complete proof of the safety theorem.

@@

\paragraph{Organization}

Throughout this report, we present in parallel our general formalized framework (with particular emphasis on differences with Alur's presentation), and the way it is used to obtain a safety proof for an example hybrid system, taken from \cite{alur}, modeling the operation of a thermostat. In section \ref{concrete}, we begin by defining the structure of normal (concrete) hybrid systems, their semantics, reachability, and safety. We also define the concrete thermostat and the safety condition we wish to prove. In section \ref{dn}, we take a moment to discuss some of the limits on decidability of properties of computable real numbers, as these complicate matters in the remainder of the development. Next, in section \ref{estimation} we describe notions of underestimation and overestimation that are applied throughout the development. Then, in section \ref{abstraction}, we begin our presentation of the abstraction method by discussing abstract regions, states, and spaces. We build on these in section \ref{abs.reach}, where we first develop the ``obvious'' notions of abstract system and reachability, discuss why they fail, and then rework them to embrace reachability sharing. In section \ref{underestimatingsafety} we show how, assuming abstract reachability is decidable, we may prove hybrid system safety by computation. Then, in section \ref{graphs}, we show how an abstract system consisting of transition and initiality ``overestimators'' can be used to decide abstract reachability by construction of a directed graph corresponding to the system, and the use of an ordinary but verified graph reachability algorithm. Finally, in sections \ref{over_cont_trans} and \ref{over_disc_trans} we describe the implementation of the aforementioned overestimators. We end with conclusions in section \ref{conclusions}.

% proceed to define abstraction bla. Following that, we show that a simple (certified) graph reachability can be employed to compute reachability in the graph corresponding to the abstract system, and that the resulting reachability judgement can be translated back through the abstract system all the way to the concrete system. We end with conclusions in section \ref{conclusions}.

\section{(Concrete) Hybrid Systems}
\label{concrete}

@@

In \cite{hintsinunification}, Asperti et al. use a similar technique (but implemented using unification hints instead of type classes) to automate reflection of algebraic expressions in concrete syntax into a an inductively defined abstract syntax.

\section{Abstraction: Regions, States, and Spaces}
\label{abstraction}

>>>>

\documentclass[runningheads]{llncs}
\ifx\pdftexversion\undefined
  \usepackage[dvips]{graphics}
\else
  \usepackage[pdftex]{graphics}
\fi
\usepackage{amsmath}
%include format.tex
%\usepackage{rotating}
\usepackage{graphicx}
\input{diagrams}
\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}
\newcommand{\eelis}{Eelis van der Weegen$^{\dagger}$}
\newcommand{\herman}{Herman Geuvers$^{\dagger\ddagger}$}
\newcommand{\dan}{Dan Synek$^{\dagger}$}
\newcommand{\adam}{Adam Koprowski$^{\P}$}

% For the final version:
%\renewcommand{\todo}[1]{}
\newcommand{\ADAM}[1] {\todo{\textsc{Adam says}: #1}}
\newcommand{\HERMAN}[1] {\todo{\textsc{Herman says}: #1}}
\newcommand{\DAN}[1] {\todo{\textsc{Dan says}: #1}}
\newcommand{\EELIS}[1] {\todo{\textsc{Eelis says}: #1}}

\newcommand{\Exists}[2]{\exists_{#1} \, . \, #2}
\newcommand{\Forall}[2]{\forall_{#1} \, . \, #2}

\def\phi{\varphi}
\newcommand{\weg}[1]{}

\newcommand{\DN}{{\sf DN}}
\newcommand{\IR}{{\mathbf R}}
\newcommand{\Heat}{{\sf Heat}}
\newcommand{\Cool}{{\sf Cool}}
\newcommand{\Chec}{{\sf Check}}
\newcommand{\Off}{{\rm Off}} 
\newcommand{\Loc}{{\rm Loc}}
\newcommand{\AState}{{\rm AState}}
\newcommand{\State}{{\rm State}}
\newcommand{\Inv}{{\rm Inv}}
\newcommand{\Init}{{\rm Init}}
\newcommand{\Region}{{\rm Region}}
\newcommand{\List}{{\rm List}}
\newcommand{\Unsafe}{{\rm Unsafe}}
\newcommand{\Safe}{{\rm Safe}}
\newcommand{\AUnsafe}{{\rm AUnsafe}}
\newcommand{\ASafe}{{\rm ASafe}}
\newcommand{\Reach}{{\rm Reach}}
\newcommand{\AReach}{{\rm AReach}}
\newcommand{\guard}{{\rm guard}}
\newcommand{\reset}{{\rm reset}}
\newcommand{\trans}{\rightarrow_{CD}}
\newcommand{\doubleheadrightarrow}{\mathrel{\rightarrow\!\!\!\!\rightarrow}}
\newcommand{\ttrans}{\doubleheadrightarrow_{CD}}
\newcommand{\attrans}{\stackrel{A}{\doubleheadrightarrow_{CD}}}
\newcommand{\contrans}{\rightarrow_C}
\newcommand{\distrans}{\rightarrow_D}
\newcommand{\acontrans}{\stackrel{A}{\rightarrow_C}}
\newcommand{\adistrans}{\stackrel{A}{\rightarrow_D}}
\newcommand{\Dur}{{\IR_{\geq 0}}}
\newcommand{\leltdec}{{\tt leltdec}}
\newcommand{\leltdecoverlap}{{\tt leltdecoverlap}}
\newcommand{\leltdecDN}{{\tt leltdecDN}}

\newcommand{\overcontrans}{{\rm over\_cont\_trans}}
\newcommand{\overdistrans}{{\rm over\_disc\_trans}}
\usepackage{color}
\usepackage{amsmath}

@@

@@

%\arrayhs % i.e. no page breaks in code blocks

@@

\author{Herman Geuvers\inst{1,2}, Adam Koprowski\inst{3}, Dan Synek\inst{1}, \\ Eelis van der Weegen\inst{1}}
\authorrunning{H. Geuvers, A. Koprowski, D. Synek, E. van der Weegen}

\institute{
Institute for Computing and Information Sciences\\ 
Radboud University Nijmegen 
\and
Faculty of Mathematics and Computer Science\\ 
Technical University Eindhoven
\and
MLState, Paris
}

\usepackage[colorlinks,pdftitle=\thetitle,pdfauthor=]{hyperref}

@@

\begin{abstract}
We have developed a hybrid system safety prover, implemented in Coq
using the abstraction method introduced by Alur, Dang and Ivan\v ci\'c
(2006).  The development includes: a formalization of the structure of
hybrid systems; a set of utilities for the construction of an abstract
system (consisting of decidable ``overestimators'' of abstract
transitions and initiality) faithfully representing a (concrete)
hybrid system; a translation of abstract systems to graphs enabling
the decision of abstract state reachability using a certified graph
reachability algorithm; a proof of the safety of an example hybrid
system generated using this tool stack.  The development critically
relies on the computable real number implementation part of the CoRN
library of formalized constructive mathematics in Coq.  The obtained
safety proofs are fully certified and do not use floating point real
numbers, but arbitrary precise approximations of computable real
numbers. It also uses a nice interplay between constructive and
classical logic via the double negation monad.

\weg{ This technical report documents our development of a hybrid
  system safety prover, implemented in Coq using the abstraction
  method introduced by Alur in \cite{alur}. The development includes:
  a formalization of the structure of hybrid systems; a systematic
  approach and generic set of support utilities for the construction
  of an abstract system (consisting of decidable ``overestimators'' of
  abstract transitions and initiality) faithfully representing a
  (concrete) hybrid system; a translation of abstract systems to
  graphs enabling decision of abstract state reachability using a
  certified graph reachability algorithm; a proof of an example hybrid
  system (taken from \cite{alur}) generated using this tool stack. The
  development critically relies on the computable real number
  implementation part of the CoRN library of formalized constructive
  mathematics. \todo{Hm, anything else worth mentioning?}  }

@@

%\setcounter{tocdepth}{2}

%\tableofcontents
%\todo{How do I get rid of that silly author index at the top of the TOC?!}

@@

In \cite{alur}, Alur et al.\ present an automated method for hybrid system
safety verification based on the construction of an \emph{abstract}
hybrid system (essentially a finite automaton) corresponding to the
hybrid system of interest. The abstract system is constructed such
that traces in the original system are represented in the abstract
system. Consequently, one can draw conclusions about reachability of
states in the concrete system from analysis of state reachability in
the abstract system. Since the abstract system is an entirely finite
discrete object (unlike the concrete system), reachability can simply
be computed using any graph reachability algorithm. Thus, the
abstraction method brings the safety verification problem from a
continuous and infinite domain into a discrete and finite domain,
where the problem can be dealt with using standard graph algorithms.
%where it is amenable to what amounts to brute force.

The prototype implementation described in \cite{alur} was developed in
a conventional programming language using libraries that use ordinary
floating point arithmetic. The potential for bugs and floating point
artifacts inherent in this approach limits the confidence one can
justifiably have in safety judgments made by such an
implementation. In this paper, we describe a reimplementation of the
basic technique of Alur et al.\ in Coq, a proof assistant based on a
rich type theory that also functions as a programming language,
letting one develop ``certified'' programs: programs whose correctness
is proved formally within the system. In our case, the program will be
one that produces machine-checked proofs of hybrid system safety,
obtained using the abstraction method.

Our development relies crucially on computations involving real
numbers, for which we use the computable real number implementation
developed by Russell O'Connor \cite{oconnor}, and part of the CoRN
library of formalized constructive mathematics \cite{corn}. Indeed,
this development showcases its use in a concrete and practical
application. Subtleties regarding the nature of these computable reals
are reflected in this development in several ways, which we will
discuss in some detail.

\section{Motivation}
The motivation for this work is threefold.

First, we want to avoid the use of floating point number computation,
because rounding errors may affect the correctness of the safety
verification that has been performed. Instead, we want to use exact
real arithmetic, which deals with infinitely precise real numbers and
more importantly, infinitely precise algorithms for computing with
real numbers. This means that, e.g.\ an algorithm for the function
$f(x) = e^x$ can produce arbitrary many output precision (``digits'')
by asking for more input precision. So, using exact real number
algorithms, we can decide $\cos(\sin(e^2))<\cos(\sin(\frac{1}{e^2}))$
by computing better and better approximations of the value of both
expressions and observing at a certain point that the inequality
holds. (Note that with this method we will never be able to determine
in finite time that two such expressions denote the same number; for
that we need a symbolic proof.) In Nijmegen we have developed a
verified library of exact real arithmetic, formalized in the proof
assistant Coq, as part of CoRN \cite{corn,oconnor}, our Constructice
Coq Repository at Nijmegen. The algorithms (for $e^x$, $\cos$, $\sin$
etc.) have been written in the programming language of Coq and have
been proved correct using its logic. Coq is sufficiently fast that the
algorithms can also be executed in reasonable time within the system,
as can be witnessed by our online certified calculator
\url{http://prover.cs.ru.nl/calc.html}. So, from this perspective,
this work can also be seen as a case study in the use of this formal
repository of exact real arithmetic: Can we use its computational
capacities to verify safety of hybrid systems? Can we use exact real
arithmetic to model the environment and verify the correctness of the
controller?

A second related point is that hybrid systems include continuous
values. Some of these are controlled by the system itself, like a
clock that can be reset, a valve that can be opened or a gas handle
that can be opened. Others represent characteristics of the
environment and can only be observed, like speed, pressure and
temperature. Due to the fact that we deal with physical systems,
measurements are always imprecise, and setting a valve in a certain
position %may on a continuous scale 
is also imprecise. 
%Of course, we don't know exactly how imprecise, and we may assume
%that better measuring instruments may arise
This implies that when a controller makes a case distinction on the
basis of $T>20$ or $T\leq 20$ -- e.g.\ if the temperature is larger
that $20$, turn the heater off, if it is $20$ or below, keep it on --
is always only ``approximately'' effective: The actual heater may
remain on until just above $20$ degrees, because of a measuring
imprecision. In a verification, one should actually take these things
into account and we believe that constructive analysis provides the
right way to model and verify hybrid systems. In constructive
analysis, we do not have (for $x,y\in\IR$) $x>y \vee x\leq y$, so we
cannot define an algorithm on the reals by making such a case
distinction. On the other hand, we {\em do\/} have, for all
$\epsilon>0$, $x+\epsilon>y \vee x-\epsilon< y$, which means that we
can get an $\epsilon$-close enough approximation to a decision between
$x>y$ and $x\leq y$.  So, the present work can also be seen as a
confirmation of our belief that constructive analysis is the
proper setting to model and verify hybrid systems.

As a third point, the method to verify safety of hybrid systems of
\cite{alur} uses so called ``predicate abstraction''. We will detail
this approach in the next section; for now it suffices to say that it
involves mapping the hybrid system to another, finite transition
system, which is then verified. However, this abstraction itself is
not verified. In this formalization we want to also verify the
abstraction itself: show that the abstract system models the concrete
one.

@@

 % In our case, the specification of the program we write is this:
 % given a description of a concrete system, the program . The whole
 % implementation, including the computational part, serves only to
 % prove a single theorem. Hence, one essentially only needs to trust
 % Coq to be convinced of the safety proofs created by our
 % development.

@@

%Wherever we say ``|CR|'', we are really referring to the type
%|CRplain| of computable reals, part of C-CoRN.

%The computational steps become part of the complete proof of the
%safety theorem.

@@

\weg{
\paragraph{Organization}

Throughout this report, we present in parallel our general formalized
framework (with particular emphasis on differences with Alur's
presentation), and the way it is used to obtain a safety proof for an
example hybrid system, taken from \cite{alur}, modeling the operation
of a thermostat. In section \ref{concrete}, we begin by defining the
structure of normal (concrete) hybrid systems, their semantics,
reachability, and safety. We also define the concrete thermostat and
the safety condition we wish to prove. In section \ref{dn}, we take a
moment to discuss some of the limits on decidability of properties of
computable real numbers, as these complicate matters in the remainder
of the development. Next, in section \ref{estimation} we describe
notions of underestimation and overestimation that are applied
throughout the development. Then, in section \ref{abstraction}, we
begin our presentation of the abstraction method by discussing
abstract regions, states, and spaces. We build on these in section
\ref{abs.reach}, where we first develop the ``obvious'' notions of
abstract system and reachability, discuss why they fail, and then
rework them to embrace reachability sharing. In section
\ref{underestimatingsafety} we show how, assuming abstract
reachability is decidable, we may prove hybrid system safety by
computation. Then, in section \ref{graphs}, we show how an abstract
system consisting of transition and initiality ``overestimators'' can
be used to decide abstract reachability by construction of a directed
graph corresponding to the system, and the use of an ordinary but
verified graph reachability algorithm. Finally, in sections
\ref{over_cont_trans} and \ref{over_disc_trans} we describe the
implementation of the aforementioned overestimators. We end with
conclusions in section \ref{conclusions}.
}

% proceed to define abstraction bla. Following that, we show that a
% simple (certified) graph reachability can be employed to compute
% reachability in the graph corresponding to the abstract system, and
% that the resulting reachability judgement can be translated back
% through the abstract system all the way to the concrete system. We
% end with conclusions in section \ref{conclusions}.

\section{Hybrid Systems and the Abstraction method}
\label{sec:hybsys}
We briefly introduce the notion of a {\em linear hybrid system}, as
used by \cite{alur}, and we show how the abstraction method is used in
that paper in order to verify safety. There are various notions
(formal definitions) of hybrid systems
\cite{henziger,lynchvaandrager}. To illustrate the definition, we give
the example of a thermostat, which is the same as \cite{alur} uses as
a running example to illustrate the abstraction method. We will use
the same example to illustrate our formal verification in Coq.

\begin{center}
\begin{figure}[htb!]
%\begin{turn}{180}
\begin{center}
%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
    \includegraphics[scale=.5]{FIGS/ThermoFig}
\end{center}
%\end{turn}
\caption{The Thermostat as an example of a Hybrid Systems \label{fig:thermostat}}
\end{figure}
\end{center}

A hybrid system is a model of how a software system, described as a
finite set of \emph{locations}, $\Loc$ 
%\footnote{We reserve the term ``state''
%  for something else.} 
with (discrete) transitions between them, acts
on and responds to a set of continuous variables (called the
\emph{continuous state space}), typically representing physical
properties of some environment (such as temperature and pressure).

In \cite{alur}, Alur requires that the continuous state space be a
convex polyhedron in $\IR^ n$. 
%In our definition of a hybrid system,
%we initially let the continuous state space be an arbitrary
%(constructive) setoid, called |Point|. We also explicitly require
%decidability of location equality.
In the running example of Figure \ref{fig:thermostat},
the discrete component consists of three locations (called $\Heat$, $\Cool$,
and $\Chec$), whose continuous state space is $\IR^2$,
where $\IR$ are the \emph{computable} reals as defined in CoRN. 
The $\Heat$ and $\Cool$ locations represent states in which the
thermostat heats and cools the environment it operates in,
respectively. The $\Chec$ state is a self-diagnostic state in which
the thermostat does not heat or cool. The first $\IR$ component in the
continuous state space represents an internally resettable clock,
while the second represents the temperature in the environment in
which the thermostat operates.

A \emph{State} in a hybrid system is a pair $(l, p)$ consisting of a
location $l$ and a point $p$ in the continuous state space
(representing the value of the continuous variables). We use $\State$
for the set of states of a hybrid system.

The {\em invariant\/} predicate defines, for each location, a set of
permitted values for the continuous variables. 
\weg{
We will use this in the
definitions of transition relations in a moment. The morphism field
expresses that this invariant respects |Point|'s setoid equality. We
further require that it is \emph{stable}, where |stable P| is defined
as |~ ~P -> P|. We discuss the reason for this stability requirement
in section \ref{dn}.
}
For the thermostat, the invariant is as follows: $$\Inv_{\Heat}(c,T) := T\leq 10\wedge c\leq 3,\quad \Inv_{\Cool}(C,t) := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$

A hybrid system also has a set of \emph{initial states}, for which we
will use the predicate $\Init(s)$.  For the thermostat example,
$\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$
\weg{
\begin{code}
  {s: State pipe location s = Heat /\ 5 <= temp s <= 10 /\ clock s == 0}
\end{code}
where |location|, |temp|, and |clock| are the obvious projections.
} 
There is the obvious requirement that the invariant holds at each
initial state.

The remaining parts of a hybrid system (i.e.\ {\em flow}, {\em guard}
and {\em reset}) describe transitions between states, which, together
with the set of initial states, determine the set of \emph{reachable}
states, representing the possible behaviors exhibited by a
hypothetical real-world implementation of the hybrid system (as
software running on a device with sensors and actuators).

%\subsection{Flow and Continuous Transitions}
%\label{concreteflow}

Each location in a concrete system has an accompanying \emph{flow
  function} which describes how the continuous variables change over
time while the system is in that location. The idea is that then
different locations correspond to different uses of actuators
available to the software system, the effects of which are described
by the flow function. For instance, in the thermostat, the flow
function corresponding to the $\Heat$ location will have the
temperature increase with time, modeling the effect of the heater
component in the imagined thermostat device.

In the canonical definition of hybrid systems, flow functions are
specified as solutions to differential equations describing the
dynamics of the continuous variables. We follow \cite{alur} in
abstracting from these, taking instead the solutions of these
differential equations, which are flow functions $\Phi$ which satisfy:
\begin{eqnarray*}
  \Phi( p, 0) &=& p \\
  \Phi (p, t + t') &=& \Phi (\Phi( p, t), t')
\end{eqnarray*}
\weg{
We further require that $\Phi$ is a morphism respecting |Point|'s
setoid equality. We bundle functions with these properties as a record
type called |Flow| (not shown), which is parametrized over the
continuous state space type, and is equipped with an implicit coercion
to the |phi| function it contains.
\todo{Unlike Alur, we don't have ``input'' (why?).}
}

We now say that there is a (concrete) \emph{continuous transition}
from a state $(l, p)$ to a state $(l', p')$ if $l = l'$ and there is a
non-negative duration $d$ such that $p' = \Phi_l( p,d)$ with the
invariant for $l$ holding at every point along the way:
$$ (l,p) \contrans (l',p') := l=l' \wedge \Exists{d\in
\Dur}{\Phi_l(p, d) = p' \wedge \Forall{t}{0 \leq t\leq d \rightarrow
\Inv_l(\Phi_l (p, t))}}.$$

In case we are in $\IR ^2$, the flow function can be expressed as the
product of two flow functions: $\Phi_l(c,T,t) = (\phi_{l,1}(c,T,t),
\phi_{l,2}(c,T,t))$. In the thermostat example, as in many other
examples of hybrid systems, $\phi_{l,1}(c,T,t)$ does not depend on $T$
and $\phi_{l,2}(c,T)$ does not depend on $c$. We call this feature
{\em separability\/} of the flow function; our development currently
relies heavily on this property. Separability makes the form of the
flow functions simpler:
$$\Phi_l(c,T,t) = (\phi_{l,1}(c,t), \phi_{l,2}(T,t))$$

In the thermostat, for all locations, $\phi_{l,1}(c,d) = c + d$ (for
the clock) and $\phi_{\Heat,2}(T,d)= T + 2 * d$,
$\phi_{\\Cool,2}(T,d)= T * e^{-d}$, $\phi_{\Chec,2}(T,d)= T * e^{-
  \frac{1}{2}d}$.

$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge \reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$

\weg{
\begin{code}
  Definition disc_trans: relation State := fun (l, p) (l', p') =>
    guard (l, p) l' /\ reset l l' p = p' /\
    invariant (l, p) /\ invariant (l', p').
\end{code}
}

It will be obvious from Figure \ref{fig:thermostat} what the guards
and reset functions are. The reset function resets the clock for all
but one of these transitions, and leaves the temperature variable as
is. Here we can discern a conceptual distinction between continuous
variables directly controlled by the system (such as the clock in the
thermostat), and variables that model a physical phenomenon and that
can only be observed (such as the temperature in the thermostat). This
distinction is not made explicit in the definition of a hybrid system;
in principle, nothing is stopping the thermostat from treating
temperature as a variable of the former kind and resetting its
value. However, this would simply make the system unimplementable.

\weg{
\begin{code}
Definition thermo_reset (l l': Location) (p: Point): Point :=
  ( match l, l' with
    | Cool, Heat | Heat, Check | Check, Heat => 0
    | _, _ => fst p
    end
  , snd p).
\end{code}
}

%\subsection{Traces and Safety}

A transition is either continuous or discrete $\trans:= \distrans \cup \contrans$. We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$:
%format trans_refl_closure (x) = x "^{*}"
$$\Reach(s):=
    \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$
Here, $\ttrans$ is the transitive reflexive closure of $\trans$.

As mentioned before, the set of reachable states represents the
possible behaviors exhibited by a hypothetical real-world
implementation of the hybrid system (as software running on a device
with sensors and actuators).

The idea now is that the purpose of a hybrid system is typically to
keep the continuous variables within certain limits. In other words,
to limit the set of reachable states to some ``safe'' subset of the
complete state space. For the thermostat example, the intent is to
keep the temperature above $4.5$ degrees at all times, and so the safe
states are defined to be those in which we have $T > 4.5$: $$\Safe(l,c,T) := T >4.5 \quad and \quad \Unsafe(l,c,T) := T\leq 4.5.$$
The goal, then, is to verify that the reachable states are all safe.

\subsection{The Abstraction Method}
There are uncountably many traces in a hybrid system, so safety is
undecidable in general. In concrete situations, safety may be (easily)
provable, however, for example, if one finds the proper {\em proof
  invariant}. As proof-invariants are often hard to find, one would
like to have more automated heuristic methods for establishing
safety. This is what the {\em predicate abstraction method\/} of
\cite{alur} does. In general one divides the continuous state space
into a finite number of convex subsets (polygons), $A_1, \ldots, A_n$,
which yields a finite {\em abstract state space}, $\AState := \{ (l,
A_i) \mid l \in \Loc, 1\leq i \leq n \}$, with an obvious embedding
$A: \State \to \AState$ of concrete states into abstract states. On this
abstract state space, one immediately defines {\em abstract continuous
  transitions\/} and {\em abstract discrete transitions\/} as follows.
\begin{eqnarray*}
(l,P) \acontrans (l',Q) &:=& l=l' \wedge \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
(l,P) \adistrans (l',Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \distrans (l,q)}.
\end{eqnarray*}
Now {\em abstract reachability\/} can be defined in the obvious way as
follows $\AReach(a) := \Exists{s_0\in\State}{\Init(s_0) \wedge A(s_0)
\attrans a}$, and also the predicates $\ASafe$ and $\AUnsafe$, stating
when abstract states are safe / unsafe are defined as expected.

This yields a finite transition system (where it may be undecidable if
there is a transition between two states!) which has the
obvious property indicated in Figure \ref{diag:abstraction}.
\begin{figure}[htb!]
$$\begin{diagram}
 s & \rArr_{CD} & s'&& s_0& \rArr_{CD} & s_1& \rArr_{CD} & s_2& \rArr_{CD} & \ldots\\
\dTo &         &\dTo &&\dTo & &\dTo &&\dTo & &\\
A(s)& \rArr^A_{CD} & A(s') && A(s_0)& \rArr^A_{CD} & A(s_1)& \rArr^A_{CD} & A(s_2)& \rArr^A_{CD} & \ldots\\
\end{diagram}$$
\caption{The abstraction function \label{diag:abstraction}}
\end{figure}
With this property, one can conclude safety of the concrete system
from safety of the abstract system: $$\text{if}\ \Forall{a\in \AState}{\AReach(a) \rightarrow \ASafe(a)}, \text{then}\ \Forall{s\in \State}{
\Reach(s) \rightarrow \Safe(s)}.$$

The interest and power of the abstraction method lies in two
facts. First, we do not need the {\em exact\/} definitions of
$\acontrans$ and $\adistrans$ to conclude safety of the concrete
system from safety of the abstract system. We only need the property
of diagram \ref{diag:abstraction}, so we can {\em overestimate\/}
$\acontrans$ and $\adistrans$ (that is: replace it with a transition
relation that allows more transitions). Second, there is a fairly good
heuristic for making the abstraction (how to divide the continuous
state space into regions) and how to decide whether there should be an
abstract transition from one abstract state to another.
This is indicated in figure \ref{fig:abstraction}.

\begin{center}
\begin{figure}[htb!]
%\begin{turn}{180}
%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
    \includegraphics[scale=.5]{FIGS/traject}
    \includegraphics[scale=.4]{FIGS/AThermo-new3}
%\end{turn}
\caption{The abstraction function computed
\label{fig:abstraction}}
\end{figure}
\end{center}

On the left side we schematically indicate how one can try to
``follow'' the path that region $A$ takes when time passes and verify
whether this path intersects with region $B$. The idea of
overestimation is that one should only {\em not\/} put an arrow $A
\acontrans B$ if we positively know that the path of $A$ will avoid
$B$, which is the case for example for paths that follow the dashed lines.

On the right side we see the abstract state space indicated for the
location $\Heat$. The abstract state space consists of rectangles,
possibly degenerated (extending to $-\infty$ or $+\infty$). For the
bounds of these rectangles one takes the values that occur in the
specification of the hybrid system (see Figure
\ref{fig:thermostat}). According to \cite{alur}, this is a good
choice. (In case one cannot prove safety in this, there is of course
the opportunity for the user to refine it.)  The gray area indicates
that from these states also abstract discrete transitions are
possible. The dashed area is unreachable, because of the invariant for
the $\Heat$ location %prevents these states from being reachable.  
All the abstract
transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
    d$, these are all the possible abstract transitions.

There is a slight subtlety with the abstract transitions that we want
to consider. If we just take $\attrans$ to be the reflexive transitive
closure of $\acontrans \cup \adistrans$, we get far too many abstract
traces. (There is no problem with the correctness of the method --
Figure \ref{diag:abstraction} still applies -- but we get way too many
traces to be able to prove any system safe.) The problem lies in the
fact that there is no transition from any point in $[0.5,1)
  \times[5,6)$ to $[0.5,1) \times[9,10)$, but the transitive closure
        of $\adistrans$ gives us such a transition anyway. 
% \ADAM{I will try to add some more explanation}
Using this method, \cite{alur} proves the correctness of the thermostat.

\weg{
\section{Motivation}

1. Verify the correctness of the abstraction itself

2. Avoid Floating Point arithmetic

3. Avoid correctness proofs of systems that are ``almost wrong''.
   (Add one arrow to Thermostat!)

- One cannot decide x>0 of a measured value
- One can't define a function on the basis of distinguishing x<=y \/ x >y
}

\section{Formalization}
We now describe the formalization in Coq and the design choices
made. Some of the design choices were due to specifics in Coq or its
type theory CiC, while other are due to our choice to work with exact
real arithmetic and a constructive logic. We will mostly discuss the
latter design choices, because we feel they are most relevant.

\subsection{(Concrete) Hybrid Systems}
\label{concrete}

@@

In \cite{hintsinunification}, Asperti et al. use a similar technique
(but implemented using unification hints instead of type classes) to
automate reflection of algebraic expressions in concrete syntax into a
an inductively defined abstract syntax.
}

\weg{
\section{Abstraction: Regions, States, and Spaces}
